Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-04-04 overall.tex
Notes clean-up
Clean up

% two dimensions: (i) backfill utilization by the PanDA Brokers; and (ii) the
% amount of runtime spent performing detector simulations.

% We based our characterizations on measuring the number of cores utilized by
% ATLAS on Titan, the overall backfill availability, the number of detector
% simulations performed on all the resources available to ATLAS and on Titan,
% and the failure rate of these simulations. We also consider the the time taken
% to setup AthenaMP and to simulate events in the ATLAS detector with it. All
% the measurements were performed between January 2016 and February 2017,
% hereafter called `experiment time window'.

% \mtnote{TODO: expand the previous paragraph to include dimensions of AthenaMP
% characterization.}

@@

% Figure~\ref{fig:backfill-utilization} (gray bars) shows the number of
% core-hours used by ATLAS during the experiment time window. ATLAS consumed a
% total of 73.8M core-hours, for an average of $\approx$7M core-hours a month,
% with a minimum of 3.3M core-hours in April 2016 and a maximum 14.8M core-hours
% in February 2017.

@@

Figure~\ref{fig:backfill-utilization} shows the total backfill availability on
Titan (gray bars) and the number of core-hours of that availability used by
ATLAS (blue bars) during the experiment time window. ATLAS consumed a total of
51.4M core-hours, for an average of $\approx$3.7M core-hours a month, with a
minimum of 1.7M core-hours in April 2016 and a maximum 7.9M core-hours in
February 2017.

\begin{figure}[htp]
    \includegraphics[clip,width=\columnwidth]{figures/backfill_consumption.pdf}

    \caption{Titan's total backfill availability: CPU core-hours (gray) and CPU
    core-hours used by ATLAS (blue). GPU core-hours unaccounted for as they
    cannot be used by ATLAS. Efficiency of PanDA Brokers defined as percentage
    of total Titan's backfill availability used by ATLAS (Red labels).}

\label{fig:backfill-utilization}
\end{figure}

PanDA Brokers' efficiency (Figure~\ref{fig:backfill-utilization}, red labels) is
defined as the fraction of core-hours utilized by the PanDA Brokers to  Titan’s
total backfill availability during the experiment time window. The brokers
reached 18\% average efficiency, with a minimum 7.8\% efficiency on May 2016 and
a maximum 30.9\% efficiency on February 2017.
% The total backfill available was 38.1M Titan core-hours in April 2016, and
% 33.1M Titan core-hours in February 2017. This shows that the measured
% difference in efficiency did not depend on a comparable difference in total
% backfill availability.
The total backfill availability was 20.3M in April 2016, and 17.6M in February
2017. This shows that the measured difference in efficiency did not depend on a
comparable difference in the total backfill availability.

% \mtnote{Elaborate further and put it at the end, after giving several of the
% measurements we would need. ``Currently, this measure of efficiency is being
% refined on the base of the actual minimum amount of walltime that PanDA Broker
% can request when submitting job requests. This is not a fixed value as it
% depends on variable initialization and configuration overheads, simulation
% time, and number of nodes obtained. Part of the undergoing effort is to
% enable the measurements of  the distribution of these parameters to derive an
% optimization function.''}

% \jhanote{I would plot the efficiency on the Y2 axis for both
% [diagrams].}\mtnote{Done but then the first diagram ended up plotting a subset
% of the data of the second diagram. I used only the second diagram.} \jhanote{I
% would suggest similar style of X-axis as previous figure. Difficult to parse
% which is April if we are going to call out a specific month. Also consistency
% in style is generally good to keep cognitive burden low.}\mtnote{Done.}

% Figure~\ref{fig:hpc-workload-utilization} shows that d

During the experiment time window, about 2.25M detector simulation jobs were
completed on Titan, for a total of 225M events processed. This is equivalent to
0.9\% of all the 250M detector simulations performed by ATLAS in the same period
of time, and 3.5\% of the 6.6B events processed by those jobs.
% Comparatively, Titan contributed 3.9\% of the total of around
% 200K\mtnote{TODO: Fix this number} cores available to ATLAS on Grid, Cloud,
% and HPC infrastructures together.
These figures confirms the relevance of supercomputers' resource contribution to
the LHC Run 2, especially when accounting for the amount of unused total
backfill availability and the improvement of PanDA efficiency across the
experiment time window.

% \begin{figure}[htp]
%     \includegraphics[clip,width=\columnwidth]{figures/cpu_hours.png}
% \caption{\mtnote{Placeholder for the diagram/data I asked to Danila and
% Sergey. Consider a table instead of a figure.}}
% \label{fig:hpc-workload-utilization}
% \end{figure}

@@

Every detector simulation executed on Titan process 100 events. This number of
events is consistent with the physics of the use case and with the average
duration of backfill availability. The duration of a detector simulation is a
function of the number of events simulated but not all events take the same time
to be simulated.
% Figure~\ref{fig:100event-distrib} shows a distribution of the simulation time
% of
1 event simulation takes from $\approx$2 to $\approx$40 minutes, with a mean of
$\approx$14 minutes. Considering that each worker node process up to 16 events
concurrently, 100 events takes an average of 105 minutes to process. As such,
PanDA brokers do not use backfill availability with less than 105 minutes
walltime.

% \begin{figure}[htp]
%     \includegraphics[clip,width=\columnwidth]{figures/tx_simulation_distribution.pdf}
%     \caption{Distribution of the time taken by a Geant4 detector simulation to
%     simulate one event. 3000 Events; 30 Titan worker nodes; 16 simulations per
%     node; 100 events per node. Average time per event $\approx$14 min. Broad
%     distribution from $\approx$2 to $\approx$40 minutes.}
% \label{fig:100event-distrib}
% \end{figure}

Figure~\ref{fig:backfill-distrib} shows the distribution of backfill
availability on Titan as a function of number of nodes and the time of their
availability (i.e., walltime). We recorded these data by polling Titan's Moab
scheduler at regular intervals during the experiment window time, while
developing and deploying PanDA Brokers. The mean number of nodes was 691, and
their mean walltime was 126 minutes. Detector simulations of 100 events, enable
to use down to 1 node for 1/2 of the mean walltime of backfill availability. As
such, it offers a good compromise for PanDA Broker efficiency.

\begin{figure*}%[htp]
    \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_wnodes_distribution.pdf}
    \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_walltime_distribution.pdf}
    \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_avail.png}
    %figures/avail_time_vs_nodes_Oct_28_Nov_3_2015.pdf}
    \caption{Backfill availability: distribution of number of work nodes (left);
    distribution of walltime in minutes (center); and a scatter plot of the two
    variables (right). 62175 measures: mean number of work nodes available 691;
    mean walltime available 126 minutes.}
\label{fig:backfill-distrib}
\end{figure*}

@@

% The number of concurrent detector simulation jobs that PanDA Broker can manage
% depends on the DTNs' resources. Each PanDA Broker can execute 1 MPI PBS job on
% up to 300 worker nodes for a total of 4,800 cores. Scaling above this
% threshold requires instantiating multiple PanDA Brokers on the same DTN and,
% once the DTN resources are saturated, on multiple DTNs.

@@

% \begin{itemize}
%     \item problem with lustre addressed with ramdisk (done)
%     \item Competition for cache in AMD processors. (done)
%     \item Comparison hardware performance Titan/Grid (data acquired)
%     \item Summary overall figure of performance analogous to those produced
%     with NGE. Diagram as discussed in F2F meeting at Rutgers. (waiting)
% \end{itemize}

@@

% AthenaMP is a multipurpose framework that needs to be configured depending on
% the type of payalod that needs to be executed. For Geant4, this configuration
% process links up to 200 libraries, all requiring filesystem read and write
% operations.

% The execution of Geant4 is mostly compute-intensive requiring to write around
% XXKB per 100 events on disk and no more than 2GB of memory.

% -----------------------------------------------------------------------------
% \subsection{PanDA Shared Library I/O Performance Impact at OLCF}

% Athena, the ATLAS framework has a configuration and initialization stage. At
% this stage, the running job is assembled on the fly from a large number of
% shared libraries. Also, at this stage, every algorithm and service is being
% configured, at run time, by a corresponding set of Python scripts, which
% results in a large number of read operations accesses to small python
% scripts, with many includes and imports Python calls.

% \mtnote{TODO: specify that this is work in progress (see Sergey's comment) and
% see how to move it to the data performance subsection}

Initially, all the shared libraries of AthenaMP and the python scripts for the
configuration stage were stored on the Spider 2 Lustre file system. However, the
I/O patterns of the initialization and configuration stages degraded the
performance of the filesystem.
% (Figure~\ref{}\mtnote{We may want to aggregate some of the diagrams about
% lustre's experiments here}).  Since Spider 2 is a center-wide file system,
% this resulted in  performance degradation for all OLCF resources and users.
% \mtnote{I am afraid the details about the trace are too specific given the
% space constraints of a SC  submission. Please feel free to uncomment it if
% you disagree.}
%
% As can be seen in Listing~\ref{mdstrace} the metadata I/O activity for ATLAS
% exhibits a spike corresponding to the beginning of the runs before tapering
% off.
%
% \begin{minipage}{\linewidth}
% \begin{lstlisting}[language=bash,frame=single,basicstyle=\ttfamily\tiny,caption=ATLAS metadata trace,label=mdstrace]
% XK7 Application 9205593
%       39012 RPCs from 300 of 300 nodes
%         ~69291.96 per sec
%           37851 LDLM_ENQUEUE RPCs    ~67229.83 per sec
%                 pmin 13us pavg 42us pmax 4983us
%           611 LDLM_CANCEL RPCs    ~1085.24 per sec
%                 pmin 10us pavg 16us pmax 32us
%           277 MDS_CLOSE RPCs    ~492.00 per sec
%                 pmin 15us pavg 20us pmax 38us
%           154 MDS_READPAGE RPCs    ~273.53 per sec
%                 pmin 170us pavg 292us pmax 671us
%           86 MDS_GETXATTR RPCs    ~152.75 per sec
%                 pmin 15us pavg 21us pmax 65us
%           30 MDS_GETATTR RPCs    ~53.29 per sec
%                 pmin 16us pavg 20us pmax 27us
%           3 MDS_REINT RPCs    ~5.33 per sec
%                 pmin 103us pavg 136us pmax 196us
%           Overall times
%                 pmin 10us pavg 42us pmax 4983us
%
% XK7 Application 9205355
%       8698 RPCs from 62 of 62 nodes
%         ~15449.13 per sec
%           8445 LDLM_ENQUEUE RPCs    ~14999.76 per sec
%                 pmin 16us pavg 41us pmax 780us
%           92 MDS_CLOSE RPCs    ~163.41 per sec
%                 pmin 15us pavg 23us pmax 52us
%           55 MDS_READPAGE RPCs    ~97.69 per sec
%                 pmin 189us pavg 291us pmax 534us
%           52 LDLM_CANCEL RPCs    ~92.36 per sec
%                 pmin 12us pavg 16us pmax 25us
%           41 MDS_GETXATTR RPCs    ~72.82 per sec
%                 pmin 16us pavg 20us pmax 27us
%           13 MDS_GETATTR RPCs    ~23.09 per sec
%                 pmin 16us pavg 21us pmax 38us
%           Overall times
%                 pmin 12us pavg 42us pmax 780us
%
% \end{lstlisting}
% \end{minipage}
%
% As this problem was identified, the OLCF staff has enabled read-only access to
% certain NFS-exported directories from Titan compute nodes. This in turn
% allowed the OLCF staff to install a software package from a Titan login node
% and have it available read-only on a Titan compute node.
%
% After careful characterization of the impact of ATLAS jobs on Lustre, the
% performance degradation
This was addressed by moving the AthenaMP distribution to a read-only NFS
directory, shared among DTNs and worker nodes. NFS eliminated the problem of
metadata contention, improving metadata read performance from $\approx$6,300
seconds on Lustre to $\approx$1,500 seconds on NFS.

% Figure~\ref{fig:atlas-perf-improvement} shows the overall ATLAS performance
% improvement on Titan. The circled region illustrates the switch from Lustre to
% NFS-exported directory for hosting the ATLAS release.

%\begin{figure}[!htb]
%    \centering
%    \begin{tabular}{cc}
%        {\includegraphics[width=0.48\textwidth]{figures/panda-completed-jobs-sw-move.pdf}}\\
%    \end{tabular}
%    \caption{ATLAS performance improvement on Titan. The circled region shows the switch from Lustre to NFS-exported directory for hosting the ATLAS release.}
%\label{fig:atlas-perf-improvement}
%\end{figure}

@@

The AMD Opteron 6274 CPU used on Titan has 16 cores, divided into 8 compute
units. Each compute units has 1 floating point (FP) scheduler shared between 2
cores. When using 16 cores for FP-intensive calculations, each pair of cores
competes for a single FP scheduler. This creates the overhead shown in
Figure~\ref{fig:comparison-8-16cores}: the mean runtime per event for 8
concurrent simulations computing 50 events is 10.8 minutes, while for 16
simulations is 14.25 minutes (consistent with the measured distribution of the
duration of event simulation). Despite an inefficiency of almost 30\%, Titan's
allocation policy based on number of worker nodes used instead of number of
cores does not justify the use of 1/2 of the cores available.

@@

The performance analysis of Titan's AMD CPUs for detector simulations helps also
to compare Titan and Grid site performances. Usually, Grid sites exposes
resources with heterogeneous CPU architectures and a maximum of 8 (virtual)
cores per worker node, while Titan's offer an homogeneous 16 cores architecture.
We used the rate of events processes per minute as a measure of the efficiency
of executing the same detector simulation on both Titan or Grid sites.
% Figure~\ref{fig:comparison-titan-grid} compares the efficiencies of Titan to
% the BNL and SIGNET Grid sites, normalized for 8 cores. Effective performance
% per-core at Titan is $\approx$0.57 event per minute, roughly 1/2 of BNL and
% 1/3 of SIGNET performances.
Comparisons of the efficiencies of Titan to  the BNL and SIGNET Grid sites,
normalized for 8 cores, show that the effective performance per-core at Titan is
$\approx$0.57 event per minute, roughly 1/2 of BNL and  1/3 of SIGNET
performances.

% \begin{figure}[htp]
%     \includegraphics[clip,width=\columnwidth]{figures/titan_grid_event_rate.pdf}
%     \caption{Comparison of the event rate for finished jobs at Titan (h1), BNL
%     (h2) and GRIDNET (h3). Titan histogram shows 1/2 of the measured values to
%     account for the core difference between Titan (16) and the Grid sites (8).
%     Performance per core: BNL 1.126; SIGNET 0.885; Titan 0.577.\mtnote{TO BE DELETED.}}
% \label{fig:comparison-titan-grid}
% \end{figure}

@@

% \mtnote{Figure summarizing job performance on Titan analogous to those
% produced with NGE, as discussed in F2F meeting at Rutgers?}
%
% ----------------------------------------------------------------------------
% \subsection{Reliability of PanDA Broker and Detector Simulations on Titan}
% \label{ssec:reliability}
%
% Figure~\ref{fig:failures-titan} shows a breakdown of the type of failures
% recorded by the PanDA Brokers between January 2016 and February 2017. The
% initial preponderance of unknown failures was progressively addressed by a
% more accurate failure model, specifically tailored to the PanDA Broker and
% Titan. Figure~\ref{fig:failures-titan} also shows the progressive improvement
% of the broker's reliability: In Februrary 2017, errors related to input files
% and initialization were almost eliminated and execution errors were comparable
% to those recorded in January 2017 even if almost twice as many simulations
% were performed (Figure~\ref{fig:backfill-utilization}). The increased amount
% of unknown failure in February indicates that furter work is needed to improve
% the PanDA Broker failure model.
%
% \begin{figure}[htp]
%     \includegraphics[clip,width=\columnwidth]{figures/failures_titan.pdf}
    % \caption{PanDA Broker failures during the experiment time window. Failures
    % were recorded with 52 exit codes, here aggregated into 6 categories. The
    % number of failures should be compared to the number of simulations
    % executed in each month as plotted in
    % Figure~\ref{fig:backfill-utilization}. Improvements in the failure model
    % of PanDA Broker shows a decrease in unknown failure and a the progressive
    % addressing of other categories of failure.}
% \label{fig:failures-titan}
% \end{figure}
%
% -----------------------------------------------------------------------------
% \subsection{Characterizing PanDA I/O}
%
% To better understand the I/O impact of ATLAS PanDA project on Titan
% supercomputing environment

We studied the impact of acquiring event descriptions on Lustre by analyzing
1,175 jobs ran on the week of 10/25/2016, for a total of 174 hours.
Table~\ref{panda-olcf-stats} shows the overall statistical breakdown of the
observed file I/O.
% Figures~\ref{fig:atlas-titan-io-read} and~\ref{fig:atlas-titan-io-written}
% show the file read and write I/O histograms for these 1,175 jobs,
% respectively. Figures~\ref{fig:atlas-titan-file-open}
% and~\ref{fig:atlas-titan-file-close} show the file $open()$ and $close()$
% metadata load histograms of the same  1,175 ATLAS jobs, respectively.
%
% As can be seen from Table~\ref{panda-olcf-stats},
%
ATLAS used between 1 and 300 worker nodes, and 35 on average. 75\% of the jobs
run by ATLAS consumed less than 25 nodes and 92\% less than 100. During the 174
hours of data collection, 6.75 ATLAS jobs were executed on average per hour,
each job running for an average of 1.74 hours.
% ATLAS jobs issues a large number of file read operations, as can be seen in
% Table~\ref{panda-olcf-stats}.
Every job read less than 250 GB and wrote less than 75 GB of data and, on
average, each job read 20 GB and wrote 6 GB of data.

@@

% \begin{figure}[!htb]
%     \centering
%     \begin{tabular}{cc}
%         {\includegraphics[width=0.48\textwidth]{figures/panda_data_read_finer_hist.pdf}}\\
%     \end{tabular}
%     \caption{ATLAS file read operation histogram on Titan for week of 10/25/16.}
% \label{fig:atlas-titan-io-read}
% \end{figure}
%
% \begin{figure}[!htb]
%     \centering
%     \begin{tabular}{cc}
%         {\includegraphics[width=0.48\textwidth]{figures/panda_data_written_finer_hist.pdf}}\\
%     \end{tabular}
%     \caption{ATLAS file write I/O histogram on Titan for week of 10/25/16.}
% \label{fig:atlas-titan-io-written}
% \end{figure}

% The I/O of the ATLAS jobs show an interesting pattern.
% Table~\ref{panda-olcf-stats} indicates that the amount of data read per ATLAS
% worker node is less than 400 MB on average, while the amount of data written
% per node is less than 170 MB on average. This correlates with our finding
% that ATLAS PanDA jobs are read heavy. However,

ATLAS jobs are read heavy: On average, the amount of data read per worker node
is less than 400 MB, while the amount of data written is less than 170 MB.
% as can be seen in Table~\ref{panda-olcf-stats} and figures
% {fig:atlas-titan-io-read} and {fig:atlas-titan-io-written},
Distributions of read and written data are different: The read operation
distribution per job shows a long tail, ranging from 12.5 GB to 250 GB, while
the written amount of data has a very narrow distribution.

% \begin{figure}[!htb]
%     \centering
%     \begin{tabular}{cc}
%         {\includegraphics[width=0.48\textwidth]{figures/panda_file_open_hist.pdf}}\\
%     \end{tabular}
%     \caption{ATLAS file $open()$ histogram on Titan for week of 10/25/16.}
% \label{fig:atlas-titan-file-open}
% \end{figure}
%
% \begin{figure}[!htb]
%     \centering
%     \begin{tabular}{cc}
%         {\includegraphics[width=0.48\textwidth]{figures/panda_file_close_hist.pdf}}\\
%     \end{tabular}
%     \caption{ATLAS file $close()$ histogram on Titan for week of 10/25/16.}
% \label{fig:atlas-titan-file-close}
% \end{figure}

The metadata I/O breakdown shows that ATLAS jobs yield 23 file $open()$
operations per second (not including file $stat()$ operations) and 5 file $close()$ operations per second, with similar distributions.
% As can be seen from figures~\ref{fig:atlas-titan-file-open}
% and~\ref{fig:atlas-titan-file-close}, they .
On average, the maximum number of file $open()$ operations per job is
$\approx$170/s and the maximum number of file $close()$ operations is
$\approx$39/s.
% on average per second per job.
For the 1,175 ATLAS jobs observed, the total number of file $open()$ operations
is 172,089,760 and the total number of file $close()$ operations is 40,132,992.
The difference between these two values is still under investigation: One
possible explanation is that ATLAS jobs don't call a file $close()$ operation
per every file $open()$ issued.

% based on our experiments with ATLAS jobs, it can be safely concluded
% that the file and metadata I/O load of ATLAS PanDA project on the OLCF Titan
% supercomputing environment and the Spider 2 file system is not detrimental to
% the center operations. At the current scale of the project, the overall impact
% of ATLAS operations on Titan is minimal.

>>>>

@@

@@

Fig.~\ref{fig:backfill-utilization} shows the total backfill availability on
Titan (gray bars) and the number of core-hours of that availability used by
ATLAS (blue bars) during the experiment time window. ATLAS consumed a total of
51.4M core-hours, for an average of $\approx$3.7M core-hours a month, with a
minimum of 1.7M core-hours in April 2016 and a maximum 7.9M core-hours in
February 2017.

\begin{figure}[htp]
    \includegraphics[clip,width=\columnwidth]{figures/backfill_consumption.pdf}
    \caption{Titan's total backfill availability: CPU core-hours (gray) and CPU
    core-hours used by ATLAS (blue). GPU core-hours unaccounted for as they
    cannot be used by ATLAS. Efficiency of PanDA Brokers defined as percentage
    of total Titan's backfill availability used by ATLAS (Red labels).}
\label{fig:backfill-utilization}
\end{figure}

PanDA Brokers' efficiency (Fig.~\ref{fig:backfill-utilization}, red labels) is
defined as the fraction of core-hours utilized by the PanDA Brokers to  Titan’s
total backfill availability during the experiment time window. The brokers
reached 18\% average efficiency, with a minimum 7.8\% efficiency on May 2016 and
a maximum 30.9\% efficiency on February 2017. The total backfill availability
was 20.3M in April 2016, and 17.6M in February 2017. This shows that the
measured difference in efficiency did not depend on a comparable difference in
the total backfill availability.

During the experiment time window, about 2.25M detector simulation jobs were
completed on Titan, for a total of 225M events processed. This is equivalent to
0.9\% of all the 250M detector simulations performed by ATLAS in the same period
of time, and 3.5\% of the 6.6B events processed by those jobs. These figures
confirms the relevance of supercomputers' resource contribution to the LHC Run
2, especially when accounting for the amount of unused total backfill
availability and the improvement of PanDA efficiency across the experiment time
window.

@@

Every detector simulation executed on Titan process 100 events. This number of
events is consistent with the physics of the use case and with the average
duration of backfill availability. The duration of a detector simulation is a
function of the number of events simulated but not all events take the same time
to be simulated. 1 event simulation takes from $\approx$2 to $\approx$40
minutes, with a mean of $\approx$14 minutes. Considering that each worker node
process up to 16 events concurrently, 100 events takes an average of 105 minutes
to process. As such, PanDA brokers do not use backfill availability with less
than 105 minutes walltime.

Fig.~\ref{fig:backfill-distrib} shows the distribution of backfill
availability on Titan as a function of number of nodes and the time of their
availability (i.e., walltime). We recorded these data by polling Titan's Moab
scheduler at regular intervals during the experiment window time, while
developing and deploying PanDA Brokers. The mean number of nodes was 691, and
their mean walltime was 126 minutes. Detector simulations of 100 events, enable
to use down to 1 node for 1/2 of the mean walltime of backfill availability. As
such, it offers a good compromise for PanDA Broker efficiency.

\begin{figure*}%[htp]
    \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_wnodes_distribution.pdf}
    \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_walltime_distribution.pdf}
    \includegraphics[clip,width=0.32\textwidth]{figures/titan_backfill_avail.png}
    \caption{Backfill availability: distribution of number of work nodes (left);
    distribution of walltime in minutes (center); and a scatter plot of the two
    variables (right). 62175 measures: mean number of work nodes available 691;
    mean walltime available 126 minutes.}
\label{fig:backfill-distrib}
\end{figure*}

@@

@@

@@

Initially, all the shared libraries of AthenaMP and the python scripts for the
configuration stage were stored on the Spider 2 Lustre file system. However, the
I/O patterns of the initialization and configuration stages degraded the
performance of the filesystem. This was addressed by moving the AthenaMP
distribution to a read-only NFS directory, shared among DTNs and worker nodes.
NFS eliminated the problem of metadata contention, improving metadata read
performance from $\approx$6,300 seconds on Lustre to $\approx$1,500 seconds on
NFS.

@@

The AMD Opteron 6274 CPU used on Titan has 16 cores, divided into 8 compute
units. Each compute units has 1 floating point (FP) scheduler shared between 2
cores. When using 16 cores for FP-intensive calculations, each pair of cores
competes for a single FP scheduler. This creates the overhead shown in
Fig.~\ref{fig:comparison-8-16cores}: the mean runtime per event for 8
concurrent simulations computing 50 events is 10.8 minutes, while for 16
simulations is 14.25 minutes (consistent with the measured distribution of the
duration of event simulation). Despite an inefficiency of almost 30\%, Titan's
allocation policy based on number of worker nodes used instead of number of
cores does not justify the use of 1/2 of the cores available.

@@

The performance analysis of Titan's AMD CPUs for detector simulations helps also
to compare Titan and Grid site performances. Usually, Grid sites exposes
resources with heterogeneous CPU architectures and a maximum of 8 (virtual)
cores per worker node, while Titan's offer an homogeneous 16 cores architecture.
We used the rate of events processes per minute as a measure of the efficiency
of executing the same detector simulation on both Titan or Grid sites.
Comparisons of the efficiencies of Titan to  the BNL and SIGNET Grid sites,
normalized for 8 cores, show that the effective performance per-core at Titan is
$\approx$0.57 event per minute, roughly 1/2 of BNL and  1/3 of SIGNET
performances.

@@

We studied the impact of acquiring event descriptions on Lustre by analyzing
1,175 jobs ran on the week of 10/25/2016, for a total of 174 hours.
Table~\ref{panda-olcf-stats} shows the overall statistical breakdown of the
observed file I/O. ATLAS used between 1 and 300 worker nodes, and 35 on average.
75\% of the jobs run by ATLAS consumed less than 25 nodes and 92\% less than
100. During the 174 hours of data collection, 6.75 ATLAS jobs were executed on
average per hour, each job running for an average of 1.74 hours. Every job read
less than 250 GB and wrote less than 75 GB of data and, on average, each job
read 20 GB and wrote 6 GB of data.

@@

ATLAS jobs are read heavy: On average, the amount of data read per worker node
is less than 400 MB, while the amount of data written is less than 170 MB.
Distributions of read and written data are different: The read operation
distribution per job shows a long tail, ranging from 12.5 GB to 250 GB, while
the written amount of data has a very narrow distribution.

The metadata I/O breakdown shows that ATLAS jobs yield 23 file $open()$
operations per second (not including file $stat()$ operations) and 5 file
$close()$ operations per second, with similar distributions. On average, the
maximum number of file $open()$ operations per job is $\approx$170/s and the
maximum number of file $close()$ operations is $\approx$39/s. For the 1,175
ATLAS jobs observed, the total number of file $open()$ operations is 172,089,760
and the total number of file $close()$ operations is 40,132,992. The difference
between these two values is still under investigation: One possible explanation
is that ATLAS jobs don't call a file $close()$ operation per every file $open()$
issued.

### 2017-04-04 workflow.tex
Clean up

\paragraph{Figure~\ref{fig:atlas_workflow} (1)} Multiple end-users (i.e.,
physicists) submit descriptions of ATLAS Monte Carlo Workflows to the JEDI
component of ATLAS middleware. JEDI supervises the execution of these workflows
through a set of eight predefined stages. Each stage requires a specific type of
input and returns a specific type of output. In this paper, we focus on the
third stage called ``Detector Simulator'': the only one, at the moment, executed
on the HPC resources of Titan.

\paragraph{Figure~\ref{fig:atlas_workflow} (2)} JEDI submits/hands over events
to a PANDA Server. This server manages the execution of all the ATLAS workflows
and their stages, including the ATLAS Monte Carlo Workflow. In
Figure~\ref{fig:atlas_workflow} we depicted only how the events of the third
stage are organized within PANDA Server. Tasks are container of jobs and each
job contains itself a set of events. Each task, each job in that task and all
the events of those jobs belongs to a specific user. This enables provenance
across the PanDA stack and isolating the management of the data of each task by
dedicated subsystems.

\paragraph{Figure~\ref{fig:atlas_workflow} (3)} Sets of events belonging to
multiple tasks and jobs are pulled from the PANDA Server by one or more PANDA
Pilot. Currently, twenty PANDA Pilots are instantiated on four dedicated DSL
nodes of Titan. These pilots are responsible for the packaging of a set of
events into a PBS job script. Packaging is performed on the base of the
information gathered about the availability of backfill resources on Titan.
Depending on the amount of nodes available on Titan, a specific number of events
is packaged into the PBS script. It should be noted that the events packaged
into a PBS job can belong to different tasks and therefore users.

\paragraph{Figure~\ref{fig:atlas_workflow} (4)} AthenaMP is used to perform the
simulation of the events within the ATLAS detector, one AthenaMP thread for each
core available on the Titan nodes. Currently, sixteen threads are used for each
node, each thread being able to simulate an average of 1 event every n??
minutes?/second? (see Figure~\ref{distribution_amp_titan}). So far, no
description has been given of the flow of data required to simulate each event.

% Questions for Danila/Sergey:
% \begin{itemize}
%     \item Does each job of a PANDA Server (and therefore all the tasks of that specific job) belong to the same `user'/VO\@?
%     \item Does a PANDA Pilot pull jobs/tasks from a PANDA Server or does a PANDA server push jobs/tasks to a PANDA Pilot?
%     \item Does a PANDA Pilot pull jobs or tasks from a PANDA Server? Are the jobs within a PANDA Pilot the same jobs (i.e., the same container of the same tasks) in the PANDA Server?
%     \item Do we have 16 PANDA Pilots on a Titan DSL (is this the right acronym?) node? How many DSL nodes?
%     \item Can the events packaged into a PBS job belong to different submissions, i.e., different jobs, i.e., different users?
%     \item Do we have the distribution of the time taken on titan to process 1 event of the Monte Carlo workflow?
% \end{itemize}

% Notes - answers
%
% - Does each job of a PANDA Server (and therefore all the tasks of that
%   specific job) belong to the same ‘user’/VO? • Does a PANDA Pilot pull
%   jobs/tasks from a PANDA Server or does a PANDA server push jobs/tasks to a
%   PANDA Pilot?
%
% Same user task->job on the server.
%
% - Does a PANDA Pilot pull jobs or tasks from a PANDA Server? Are the jobs
%   within a PANDA Pilot the same jobs (i.e., the same container of the same
%   tasks) in the PANDA Server?
%
% Panda pilot is a edge service on titan, event-driven by backfill
% availability. When available, it DOES THE PULL from the panda server.
%
% - Do we have 16 PANDA Pilots on a Titan DTN (is this the right acronym?)
%   node? How many DTN nodes?
%
% Currently 20, thinking to go higher, on 4 DTN (Data Transfer Node), 5 pilots
% for DTN, DTN that enable interactive data transfer. New DTN farm is being
% rolled on. Our set up will change. More node, more pilots per node. This is
% bound by DTN performance, overloading DTN (currently absence of data on why
% these nodes are overladed. Network/memory should not be a problem, it is
% probably purely CPU because of tools used to move files, e.g., GridFTP.)
%
% - Can the events packaged into a PBS job belong to different submissions,
%   i.e., different jobs, i.e., different.
%
% They do belong to different tasks.
%
% - Do we have the distribution of the time taken on titan to process 1 event
%   of the Monte Carlo workflow?
%
% To be added to Section VI about performance. See presentation:
% https://drive.google.com/drive/u/0/folders/0BxWZQExiTsWNRTdMQXN4a1g3bHc

>>>>

\paragraph{Fig.~\ref{fig:atlas_workflow} (1)} Multiple end-users (i.e.,
physicists) submit descriptions of ATLAS Monte Carlo Workflows to the JEDI
component of ATLAS middleware. JEDI supervises the execution of these workflows
through a set of eight predefined stages. Each stage requires a specific type of
input and returns a specific type of output. In this paper, we focus on the
third stage called ``Detector Simulator'': the only one, at the moment, executed
on the HPC resources of Titan.

\paragraph{Fig.~\ref{fig:atlas_workflow} (2)} JEDI submits/hands over events
to a PANDA Server. This server manages the execution of all the ATLAS workflows
and their stages, including the ATLAS Monte Carlo Workflow. In
Fig.~\ref{fig:atlas_workflow} we depicted only how the events of the third
stage are organized within PANDA Server. Tasks are container of jobs and each
job contains itself a set of events. Each task, each job in that task and all
the events of those jobs belongs to a specific user. This enables provenance
across the PanDA stack and isolating the management of the data of each task by
dedicated subsystems.

\paragraph{Fig.~\ref{fig:atlas_workflow} (3)} Sets of events belonging to
multiple tasks and jobs are pulled from the PANDA Server by one or more PANDA
Pilot. Currently, twenty PANDA Pilots are instantiated on four dedicated DSL
nodes of Titan. These pilots are responsible for the packaging of a set of
events into a PBS job script. Packaging is performed on the base of the
information gathered about the availability of backfill resources on Titan.
Depending on the amount of nodes available on Titan, a specific number of events
is packaged into the PBS script. It should be noted that the events packaged
into a PBS job can belong to different tasks and therefore users.

\paragraph{Fig.~\ref{fig:atlas_workflow} (4)} AthenaMP is used to perform the
simulation of the events within the ATLAS detector, one AthenaMP thread for each
core available on the Titan nodes. Currently, sixteen threads are used for each
node, each thread being able to simulate an average of 1 event every n??
minutes?/second? (see Fig.~\ref{distribution_amp_titan}). So far, no
description has been given of the flow of data required to simulate each event.
